---
title: "R Lessons Learned"
output: github_document
---

```{r}
library(ggplot2)
```

Here are some lessons I've (re)learned recently in consultation with other R users. Once you've googled the same thing several times, you tend to want to make a note of it. And [if you give a data librarian an R notebook, they'll want to post it on GitHub](https://en.wikipedia.org/wiki/If_You_Give_a_Mouse_a_Cookie). Maybe these lessons will help someone else!

## How to fix a dataframe with an unlabelled first column (convert row names into first column)

### Problem

Example:

```{r}
mtcars
names(mtcars)
```

There is no apparent way to reference the cars' model names.

### Solutions

What appears as the "first column" is really the **row names**:

```{r}
rownames(mtcars)
```

There are several ways to prepend a column made of the row names.

```{r}
new_mtcars <- cbind(make_model=rownames(mtcars), data.frame(mtcars, row.names=NULL))
new_mtcars
```

Source/more: [https://stackoverflow.com/questions/29511215/convert-row-names-into-first-column](https://stackoverflow.com/questions/29511215/convert-row-names-into-first-column)

## How to remove elements of a ggplot, e.g., axes and tickmarks

Default appearance:

```{r}
ggplot(data=mtcars, mapping=aes(x=cyl, y=mpg)) +
  geom_point()
```

Use `element_blank()` inside of a `theme()` call: (example: remove axis ticks)

```{r}
ggplot(data=mtcars, mapping=aes(x=cyl, y=mpg)) +
  geom_point() +
  theme(axis.ticks=element_blank())
```

